\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-02-24 redaction/dhp_sept2016.tex
ajout intro

\todo{reprendre slides presentation LEDa pour la motivation de la question.}

@@

Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{\textbf{donner l'intuition. Deformation des prix relatifs, TBC}} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to fill this gap. \bigskip

>>>>

%\todo{reprendre slides presentation LEDa pour la motivation de la question.}
%JH: apparently, it has been done! Or do you want to have the Italian shoes example in the intro????

@@

Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs.\footnote{This is due to the alteration of relative prices in a heterogeneous-firms trade framework. Lower additive trade costs reduce the export market prices of the highly performing exporters relative to lowly performing exporters and thus shift relative market shares among exporters towards the highly performing exporters - this of course does not happen with iceberg costs. Total export sales are then reallocated between a smaller number of high productivity firms. Therefore, the expected profits from exporting prior to entry increase all other things being equal by more in the case of lower per-unit costs, driving up incentives to enter the industry and bringing eventually the welfare enhancing intra-industry reallocations highlighted in \citet{melitz} framework.} Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized ad-valorem tariff. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to fill this gap. \bigskip
